Paper Analysis

### 2024-12-10 Paper_WriteUp/Paper_Plots.qmd
modify the helper in some places

## for each player, count how many times they start
## a game with 5 minutes of the start of a previous game
## and get the proportion across all focal players
 bullet_data |>  
   group_by(Username) |> 
   select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
   group_by(Username) |> 
   arrange(date_time) |> 
   mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
   mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
   summarise(sum(less_5, na.rm = TRUE)/ n())

@@

## repeat for blitz data using 7 minutes as the maximum gap
blitz_data |>  
   group_by(Username) |> 
   select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
   group_by(Username) |> 
   arrange(date_time) |> 
   mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
   mutate(less_7 = ifelse(gap_seconds <= 420, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
   summarise(sum(less_7, na.rm = TRUE)/ n())

## examine how many games are played in a session for the bullet data
## look at the streaks, number of games played in a session
outcomes <- bullet_data |> 
  # group_by(Username) |> 
  select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, missing = FALSE))

@@

## repeat for blitz games
outcomes_blitz <- blitz_data |> 
  select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7 = if_else(gap_seconds <= 420, TRUE, FALSE, missing = FALSE))

@@

bullet_data_gm |> nrow()

@@

## get the average rating of the GMs at the time data
## was collected 
lichess_data_gm |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated blitz game", Event)) |> 
  group_by(Username) |> 
  arrange(UTCDate, UTCTime) |> 
  slice_tail(n = 1) |> 
  mutate(user_elo = ifelse(Username == White, WhiteElo, BlackElo)) |> 
  select(Username, White, Black, WhiteElo, BlackElo, user_elo) |>
  ungroup() |> 
  summarise(mean(user_elo))

## look at the proportion of games started within 5 minutes 
## of the previous game
bullet_data_gm |> 
  select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
  ungroup() |>   ## to get the overall proportion
  summarise(sum(less_5, na.rm = TRUE)/ n())

## look at the streaks, number of games played in a session
outcomes_gm <- bullet_data_gm |> 
  # group_by(Username) |> 
  select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_5 = if_else(gap_seconds <= 300, TRUE, FALSE, missing = FALSE))

@@

## repeat for blitz
outcomes_blitz <- blitz_data_gm |> 
  select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7_min = if_else(gap_seconds <= 420, TRUE, FALSE, FALSE))

@@

perm_17_19_winner <- player_pars |> 
  filter(param == 1) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

@@

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)
```

@@

>>>>

## for each player, count how many times they start
## a game with 5 minutes of the start of a previous game
## and get the proportion across all focal players
 bullet_data |>  
   group_by(Username) |> 
   dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
   group_by(Username) |> 
   arrange(date_time) |> 
   mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
   mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
   summarise(sum(less_5, na.rm = TRUE)/ n())

@@

## repeat for blitz data using 7 minutes as the maximum gap
blitz_data |>  
   group_by(Username) |> 
   dplyr::select(Username, UTCDate, UTCTime) |> 
   mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
   group_by(Username) |> 
   arrange(date_time) |> 
   mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
   mutate(less_7 = ifelse(gap_seconds <= 420, TRUE, FALSE)) |> 
   ungroup() |>   ## to get the overall proportion
   summarise(sum(less_7, na.rm = TRUE)/ n())

## examine how many games are played in a session for the bullet data
## look at the streaks, number of games played in a session
outcomes <- bullet_data |> 
  # group_by(Username) |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_3 = if_else(gap_seconds <= 180, TRUE, FALSE, missing = FALSE))

@@

## repeat for blitz games
outcomes_blitz <- blitz_data |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7 = if_else(gap_seconds <= 420, TRUE, FALSE, missing = FALSE))

@@

bullet_data_gm |> nrow()
bullet_data_gm |> 
  group_by(Termination) |>
  tally() |> 
  mutate(prop = n/sum(n))

@@

## get the average rating of the GMs at the time data
## was collected 
lichess_data_gm |> 
  mutate(Event = tolower(Event)) |> 
  filter(grepl("rated blitz game", Event)) |> 
  group_by(Username) |> 
  arrange(UTCDate, UTCTime) |> 
  slice_tail(n = 1) |> 
  mutate(user_elo = ifelse(Username == White, WhiteElo, BlackElo)) |> 
  dplyr::select(Username, White, Black, WhiteElo, BlackElo, user_elo) |>
  ungroup() |> 
  summarise(mean(user_elo))

## look at the proportion of games started within 5 minutes 
## of the previous game
bullet_data_gm |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_5 = ifelse(gap_seconds <= 300, TRUE, FALSE)) |> 
  ungroup() |>   ## to get the overall proportion
  summarise(sum(less_5, na.rm = TRUE)/ n())

## look at the streaks, number of games played in a session
outcomes_gm <- bullet_data_gm |> 
  # group_by(Username) |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_5 = if_else(gap_seconds <= 300, TRUE, FALSE, missing = FALSE))

@@

## repeat for blitz
outcomes_blitz <- blitz_data_gm |> 
  dplyr::select(Username, UTCDate, UTCTime) |> 
  mutate(date_time = ymd_hms(paste(UTCDate, UTCTime))) |> 
  group_by(Username) |> 
  arrange(date_time) |> 
  mutate(gap = date_time - lag(date_time),
         gap_seconds = as.numeric(gap)) |> 
  mutate(less_7_min = if_else(gap_seconds <= 420, TRUE, FALSE, FALSE))

@@

perm_17_19_winner <- player_pars |> 
  filter(param == 2) |> 
  ggplot(aes(mean)) +
  geom_histogram() +
  facet_wrap(~player_name, ncol = 5, scales = "free") +
  labs(x = element_blank(), y = element_blank()) +
  geom_vline(data = true_beta |> filter(param == 2), aes(xintercept = mean),
             col = "red") +
  theme_5col()

@@

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

levels <- true_beta |> 
  filter(param == 2) |> 
  arrange(mean) |> 
  pull(player_name)

true_vals <- true_beta |> 
  filter(param == 2) |>  
  mutate(y = row_number() - 0.5,
         yend = y + 1)

perm_17_19_winner <- player_pars |> 
  filter(param == 2) |>
  mutate(y = factor(player_name, levels = levels)) |> 
  ggplot(aes(x = mean, y = y)) +
  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
  geom_segment(data = true_vals, 
               aes(x = mean, y = y, yend = yend), col = "red") +
  labs(y = "", x = "Estimated W/L-Effect") +
  theme_single_y()

ggsave(plot = perm_17_19_winner, 
       filename = paste0(here("Paper_WriteUp", "paper_figures",
                              "perm_17_19_winner.png")), dpi = 600,
       height = 5, width = 7)

@@

perm_gm_winner <- player_pars |> 
  filter(param == 2) |>
  mutate(y = factor(player_name, levels = levels)) |> 
  ggplot(aes(x = mean, y = y)) +
  stat_histinterval(slab_fill = NA, .width = c(0.5, 0.95)) +
  geom_segment(data = true_vals, 
               aes(x = mean, y = y, yend = yend), col = "red") +
  theme_single_y() +
  labs(x = "Estimated W/L-Effect", y = "")
